% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Something something Pc4 pulsations. Ultra low frequency (ULF) waves with periods of a minute or two. They correspond to field line resonances near Earth's plasmapause. Drift or drift-bounce resonance with energetic radiation belt and ring current particles. Radial diffusion. 

Pc4 pulsations are known to exhibit different behavior based on their azimuthal modenumber. Low-\azm Pc4 pulsations are driven at the outer edge of the magnetosphere, and have a compressional component. High-\azm Pc4 pulsations are non-compressional and are driven within the magnetosphere. 

Giant pulsations are a subset of high-\azm Pc4 pulsations which are of particular interest. 

High-\azm Pc4 pulsations are hard to simulate. Traditionally, simulations are driven from the outer boundary, which doesn't work in this case. And resolving high azimuthal modenumbers is computationally expensive in a 3D simulation. 

\cref{ch_intro} gives a general introduction is made to Earth's magnetosphere. (Note: this doesn't exist yet.)

\cref{ch_pc4s} summarizes Pc4 pulsations in terms of prominent theoretical and observational work, including work on giant pulsations. (Note: this mostly doesn't exist yet.)

\cref{ch_model} presents a 2.5D model designed to simulate Pc4 pulsations, including those with high azimuthal modenumbers. The model is built upon recent work by Lysak, and includes a dipole-aligned geometry, a height-resolved ionosphere, and coupling to a conducting Earth. Ring current modulation is introduced as a novel driving mechanism. 

\cref{ch_math} derives dispersion relations for a cold, resistive plasma, such as Earth's inner magnetosphere and ionosphere. This gives an idea as to the waves expected to be produced. 

\cref{ch_rbsp} compares model output to poloidal Pc4 observations made with the Van Allen Probes. (Note: this hasn't quite happened yet. The chapter briefly explains.)

\cref{ch_inertia} investigates the results of adding electron inertial effects to the model, allowing the computation of parallel electric fields and field-aligned currents. 

\cref{ch_azm} showcases changes in Pc4 behavior as a result of altering the azimuthal modenumber. This includes a rotation of poloidal to toroidal waves, significantly refining past work by Mann, and before that Radoski. (There are a lot of figures here. It might end up being two chapters.)

\cref{ch_conclusion} offers a summary and possibilities for future work. 

NOTE: ALL FIGURES ARE PDFS. Some of the frames are small, but they should remain sharp when you zoom in on them. 


