% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Acknowledgements
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This work would not have been possible without the guidance --- and patience
--- of my advisor, Bob Lysak. In hindsight, it feels like I spent these years
learning how to take ownership of a project, and that plasma physics,
differential geometry, and Fortran are just a few things I picked up on the
side. 

I would also like to thank John Wygant, Dai Lei, Ian Mann, Aaron Breneman,
Scott Thaller, and Sheng Tian, each of whom gave thoughtful responses to my
endless questions. Not once did I catch flak for being a theorist who has
never so much as set foot on a satellite. 

My sanity during the years of graduate work is a credit to the friends who
accompanied me through them, in particular, Barry Costanzi, Alex Gude, Roxanne
Radpour,
Dominick Rocco, Allan and Mandy Straub, and Jan Zirnstein. Their sanity, in
turn, is a credit to my family: my parents, Cathy and Dave; my brother, Joey;
and my partner, Kara. 

\vfill

This work was funded by the Office of the Vice President for Research at the
University of Minnesota, as well as NSF grant AGS-1405383 and NASA grant
NNX12AD14G. Supercomputer resources were provided by the Minnesota Supercomputing
Institute. Travel support was provided in part by GEM. 

%\rule{\textwidth}{1pt}

%may have never found my way into graduate school, much less out of it

%undergraduate advisor Amy Kolan, de facto co-advisor Olaf Hall-Holt
%always spoke to me with surprising candor

% Bob

% John

% MSI

% Lei, Ian

% Aaron, Scott

% Sheng. Also Adam

% Cindy, Tom, Lindsay

% Alex. Also Jan, Allan, Mandy, Roxanne

% Amy, Olaf?

% Ted, Heidi?

% Funding

% Family. Friends. Kara. 


