


\chapter{Introduction}
  \label{ch_intro}

%ALEX NOTES: 
%- Should decide if you're going to use contractions or not; many would say not to in formal writing

%\todo{In 1859, humanity was working hard to get its shit together.} 
%Austro-Sardinian War

1859 was a pivotal year in human history. The United States moved steadily 
toward the American Civil War, which would abolish slavery and consolidate the 
power of the federal government. A slew of conflicts in Southern Europe set the
stage for the unification of Italy. The Taiping Civil War --- one of the
bloodiest conflicts of all time --- is considered by many to mark the beginning
of modern Chinese history. Origin of Species was published. The first
transatlantic telegraph cable was laid.

Meanwhile, ambivalent to humanity, the Sun belched an intense burst of charged
particles and magnetic energy directly toward Earth. The resulting geomagnetic
storm\footnote{The Solar Storm of 1859 is also called the Carrington Event,
after English astronomer Richard Carrington. He drew a connection between the
storm's geomagnetic activity and the sunspots he had observed the day before.}
caused telegraph systems to fail across the Western hemisphere, electrocuting
operators and starting fires\cite{green_2006,tsurutani_2003}. Displays of the
northern lights were visible as far south as Cuba. 

The Solar Storm of 1859 was perhaps the most powerful in recorded history, but
by no means was it a one-time event. The Sun discharges hundreds of coronal
mass ejections (CMEs) per year, of all sizes, in all directions. In fact, a
comparably large CME narrowly missed Earth in 2012\cite{nasa_2012}. Had it not,
it's estimated it would have caused widespread, long-term electrical outages,
with a damage toll on the order of \num{e12} dollars\cite{lloyds_2013}. 

The Sun's extreme --- and temperamental --- effect on Earth's magnetic
environment makes a compelling case for the ongoing study of space weather.
Such research has evolved over the past century from sunspot counts and compass
readings to multi-satellite missions and supercomputer simulations. Modern
methods have dramatically increased humanity's understanding of the
relationship between the Sun and the Earth; however, significant uncertainty
continues to surround geomagnetic storms, substorms, and the various energy
transport mechanisms that make them up. 

The present work focuses in particular on the phenomenon of field line
resonance: \Alfven waves bouncing between the northern and southern
hemispheres. Such waves play an important part in the energization of
magnetospheric particles, the transport of energy from high to low altitude,
the precipitation of particles into the atmosphere, and the driving of currents
at the top of the atmosphere. It is these currents which give rise to
potentially-catastrophic magnetic disturbances at Earth's surface. 

The study of resonance in the near-Earth environment is furthermore valuable as
a proxy for other (less-accessible) plasma environments. Similar waves occur in
astrophysical plasmas, observation of which is limited by distance. Field line
resonance is also analogous to the so-called ``fishbone instability'' in fusion
reactors. As a plasma laboratory, the magnetosphere is unique in that it is
both close enough to measure directly, and also large enough that measurements
can be performed without disrupting its behavior. 

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
\section{Structure of the Present Work}

The present work is laid out as follows. 

\cref{ch_environment} surveys the near-Earth environment. Prominent features of
the magnetosphere are defined. The behavior of the magnetosphere during
geomagnetic storms and substorms is summarized. 

\cref{ch_flrs} introduces the field line resonance phenomenon, in terms of both
the underlying physics and notable work on the topic. Jargon is introduced to
clarify important elements of wave structure. Several open questions about
field line resonances (FLRs) are offered as motivations for the present work. 

\cref{ch_math} lays the ground work for a numerical model by exploring the
fundamental equations of waves in a cold, resistive plasma --- such as
Earth's magnetosphere. Characteristic scales are computed from the resulting
dispersion relations. 

\cref{ch_model} presents Tuna, a new two and a half dimensional simulation
designed specifically for the realistic modeling of FLRs. Tuna's
non-orthogonal geometry, height-resolved ionosphere, novel driving mechanism,
and coupling to the atmosphere are justified and explained. 

\cref{ch_inertia} considers the addition of electron inertial effects to Tuna,
which are neglected in the core model presented in \cref{ch_model}. These
effects allow the computation of parallel currents and electric fields, which
have not previously been included in global \Alfven models. The effects are
shown to be instability-prone and computationally expensive, but some results
are gleaned nonetheless. 

\cref{ch_results} showcases the core numerical results of the present work,
addressing several of the questions posed in \cref{ch_flrs}. The interplay
between compressional propagation, poloidal-to-toroidal rotation, and Joule
dissipation is considered from several angles. 

\cref{ch_rbsp} puts the numerical results in context through the analysis of
data from the Van Allen Probes mission. FLR occurrence rates are considered in
terms of location, harmonic, and polarization, parameters which have been only
partially addressed in past FLR surveys. 

\cref{ch_conclusion} briefly summarizes the results shown in the above chapters
--- code development, analysis of numerical results, and satellite observation
--- and suggests further directions. 

%\todo{The storm of 1859 presented compelling evidence that the Sun drives geomagnetic activity. In the decades that followed, a model took shape to describe the mechanisms of energy transfer between the Sun and the Earth. Based on auroral observations and data from ground-based magnetometers, Birkeland argued for the existence of a constant outflow of electrons and ions from the Sun -- the solar wind. The advent of high-frequency radio communication allowed Kennelly, Heaviside, and others to probe the electrical properties of the upper atmosphere. }

%\todo{The study of space weather was revolutionized by the development of sounding rockets and satellites in the mid twentieth century. This allowed direct observation of the structure of the near-Earth environment, including, crucially, the waves that carry energy through it. }

%\todo{Not least among these advances was the discovery of \Alfven waves. {\Alfven}ic aurora. Carry energy and particles. }

%The study of space weather revolves around the transfer of energy from the Sun to the Earth. Ultra low frequency waves in particular are an important energy transport mechanism between the magnetosphere's outer boundary (at the solar wind) and its inner boundary (at the top of the atmosphere). 

%\footnote{Uppercase \X, \Y, and \Z are used to indicate GSE coordinates: \Xhat points from the Earth to the Sun; \Yhat is perpendicular to \Xhat in the Sun's ecliptic plane, pointing duskwards; \Zhat points north, out of the ecliptic plane. In later chapters, lowercase \x, \y, and \z are used to define a more or less analogous corodinate system with respect to Earth. }

% The moon is at \SI{60}{\RE}. 

% Transient solar wind phenomena, such as coronal mass ejections, are also known to be related to geomagnetic disturbances at Earth. Jesse cites here: 

% R. L. McPherron. Physical processes producing magnetospheric substorms and mangetic storms. In J. A. Jacobs, editor, Geomagnetism, volume 4, chapter 7. Academic Press, 1991.

% G. Rostoker. Substorms. In Handbook of the Solar-Terrestrial Environment, chapter 15. Springer-Verlag, 2007.

% M. Shulz. Magnetospheres. In Handbook of the Solar-Terrestrial Environment, chapter 7. Springer-Verlag, 2007.

% papers mentioned during Yan's talk. mostly about alfven acceleration and nonlinear effects. 
% Vasyliunas 1970, 1984
% Hasegawa 1976
% Goertz 1991
% Stasiewicz et al 2000
% Haerendel 2008
% Song & Lysak 1994, 1999, 2000, 2001, 2006, 2011, 2012
% Inverted V?
% Double layers? 
% Charge holes? 

