
% =============================================================================
% =============================================================================
% =============================================================================
\section{Event Selection}




Note that, in gathering his 860 events, Dai\cite{dai_2015} did not take the $\vec{E} \cdot \vec{B} = 0$ assumption to clean up the electric field data. The present work has taken that assumption -- and discarded all events within \SI{15}{\degree} of the probe's spin axis. 

This cuts it down to ... events. 

Furthermore distinguish the fundamental mode from the second harmonic...

Insist on a certain confidence... 



% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
\subsection{Identifying the Fundamental Harmonic}

Takahashi\cite{takahashi_2013} spells this out explicitly. Fundamental mode, south of the equator, V should lead B by 90 degrees. 

Per Takahashi\cite{takahashi_2011}, phase lag can be used to distinguish the fundamental harmonic from the second harmonic. 

Chisham and Orr\cite{chisham_1991} argue that around \SI{7}{\RE}, frequency around \SI{10}{\mHz} precludes higher harmonics. Or maybe look at \cite{green_1985}?

Green\cite{green_1979} and Cummings\cite{cummings_1969} talk about the frequencies for ideal toroidal modes. 

Dai\cite{dai_2015} says to look at \cite{takahashi_2011} and \cite{dai_2013} for unambiguous identification of the fundamental mode. 

\cite{liu_2010} talks about setting the coordinate system using the time-averaged magnetic field, and the azimuthal direction being $\hat{B_0} \times \hat{r}$. 

